Untitled CAV Paper

### 2016-01-23 paper.tex
Abstract + some unbounded + some bounded alg + editing

@@

\begin{abstract}
    This is the abstract
\end{abstract}

@@

A \emph{safety game}, $G = \langle X, L_c, L_u, \delta, I, E, \rangle$,
consists of a set of boolean state variables, sets of controllable and
uncontrollable label variables, a transition relationship $\delta : (X, L_c,
L_u) \to X$, an initial state, and an error state. The \emph{controller} and
\emph{environment} players choose controllable and uncontrollable labels
respectively and the game proceeds according to $\delta$.

An \emph{run} of a game $(x_0, c_0, u_0), (x_1, c_1, u_1) \dots (x_n, c_n,
u_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
\leftarrow \delta(x_k, c_k, u_k)$. A run is winning for the controller if $x_0
= I \land \forall i \in \{1..n\} (x_i \neq E)$. In a bounded game of rank $n$
all runs are restricted to length $n$, whereas unbounded games consider runs of
infinite length.

A \emph{controller strategy} $\pi^c : X \to L_c$ is a mapping of states to
controllable labels. A controller strategy is winning in a bounded game of rank
$n$ if all runs $(x_0, \pi^c(x_0), u_0), (x_1, \pi^c(x_1), u_1) \dots (x_n,
\pi^c(x_n), u_n)$ are winning. Bounded \emph{realizability} is the problem of
determining the existence of such a strategy for a bounded game.

An \emph{environment strategy} $\pi^e : (X, L_c) \to L_u$ is a mapping of
states and controllable labels to uncontrollable labels. A bounded run is
winning for the environment if $x_0 = I \land \exists i \in \{1..n\} (x_i = E)$
and an environment strategy is winning for a bounded game if there exists a run
$(x_0, c_1, \pi^e(x_1, c_1)), (x_1, c_1, \pi^e(x_1, c_1)) \dots (x_n, c_n,
\pi^e(x_n, c_n))$ that wins for the environment. Safety games are zero sum,
therefore the existence of a controller strategy implies the nonexistence of an
environment strategy and vice versa.

@@

A set of runs can be symbolically represented by a tree of labels. Each edge of
the tree is either a \emph{fixed} valuation of controllable or uncontrollable
variables, or it is \emph{unfixed} denoting any values are possible. Also, the
tree must alternate between controllable and uncontrollable edges to reflect
the alternating turns of the game. The entire set of runs of a bounded game of
rank $n$ is symbolically represented by a tree of depth $2n$ and width 1
populated by unfixed edges. Reducing the set of runs in a game forms an
abstract game, which can be represented symbolically by an \emph{abstract game
tree}.

@@

@@

>>>>

\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\EndIIf}{}

@@

\begin{abstract}

    Reactive synthesis techniques based on image computations have been shown
    to work well in many cases but suffer from state explosion in others.
    A different approach applies SAT solvers in a counterexample guided framework but
    is limited to bounded synthesis.  We present an extension of this technique
    to unbounded synthesis by employing interpolation.  Experimental results
    show this to be a promising alternative that solves some previously
    intractable instances.

\end{abstract}

@@

A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E, \rangle$,
consists of a set of boolean state variables, sets of uncontrollable and
controllable label variables, a transition relationship $\delta : (X, L_u, L_c)
\to X$, an initial state, and an error state. The \emph{controller} and
\emph{environment} players choose controllable and uncontrollable labels
respectively and the game proceeds according to $\delta$.

An \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
c_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
\leftarrow \delta(x_k, u_k, c_k)$. A run is winning for the controller if $x_0
= I \land \forall i \in \{1..n\} (x_i \neq E)$. In a bounded game of rank $n$
all runs are restricted to length $n$, whereas unbounded games consider runs of
infinite length. Since we consider only deterministic games a run is equivalent
to a list of assignments to $L_c$ and $L_u$.

A \emph{controller strategy} $\pi^c : (X, L_u) \to L_c$ is a mapping of states
and uncontrollable inputs to controllable labels. A controller strategy is
winning in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)),
(x_1, u_1, \pi^c(x_1, u_1)) \dots (x_n, u_n, \pi^c(x_n, u_n))$ are winning.
Bounded \emph{realizability} is the problem of determining the existence of
such a strategy for a bounded game.

An \emph{environment strategy} $\pi^e : X \to L_u$ is a mapping of states to
uncontrollable labels. A bounded run is winning for the environment if $x_0
= I \land \exists i \in \{1..n\} (x_i = E)$ and an environment strategy is
winning for a bounded game if there exists a run $(x_0, \pi^e(x_1), c_1), (x_1,
\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ that wins for the environment.
Safety games are zero sum, therefore the existence of a controller strategy
implies the nonexistence of an environment strategy and vice versa.

@@

A set of runs can be represented symbolically by a tree of valuations to
controllable and uncontrollable label variables where valuations are allowed to
be \emph{unfixed}. An unfixed value symbolically represents all possible
valuations. Formally, we define the tree as a set of lists of label valuations
and a node in the tree may be identified by the list made by following the path
from the root.

The entire set of runs of a bounded game of rank $n$ is
symbolically represented by a tree of depth $2n$ and width 1 populated by
unfixed edges.  Reducing the set of runs in a game forms an abstract game,
which can be represented symbolically by an \emph{abstract game tree}.

@@

\begin{algorithm}
    \begin{algorithmic}
        \Function{solveAbstract}{$\langle s, r \rangle , gt$}
        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, gt$}
        \IIf{$r = n - 1$} \Return $cand$ \EndIIf 
        \State $gt' \gets gt$
        \Loop
            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf
            \State $cex \gets $ \Call{verify}{$\langle s, r \rangle, gt, cand$}
            \IIf{$cex = NULL$} \Return $cand$ \EndIIf
            \State $gt' \gets gt' \cup cex$
%%%            \State $gt' \gets \Call{refine}{gt', cex}$
            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, gt'$}
        \EndLoop
        \EndFunction
    \end{algorithmic}

%%%    \begin{algorithmic}
%%%        \Function{refine}{$gt, \langle leaf, spoiling \rangle$}
%%%            \State \Return $gt \cup \{leaf, spoiling\}$
%%%        \EndFunction
%%%    \end{algorithmic}

    \begin{algorithmic}
        \Function{findCandidate}{$\langle s, r \rangle, T$}
            \State $\hat{T} \gets \Call{gtExtend}{T}$
            \State $fml \gets $ \Call{treeFormula}{$\hat{T}$}
            \State $sol \gets $ \Call{SAT}{$s(X_T) \land fml$}
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}
        \Function{verify}{$\langle r, s\rangle, gt, cand$}
            \For{$l \in leaves(gt)$}
                \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle r, s\rangle, gt, l$}
                \State $spoiling \gets $ \Call{solveAbstract}{$\langle r', s' \rangle, \emptyset$}
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
        \node [] (root){}
            child {node [circle,draw,label=left:$x_1$] {}
                child {node [circle,draw] {}
                    child {node [circle,draw,label=left:$x_0^0$] {} edge from parent[unfixed]}
                    edge from parent [fixed] node [left] {$u_1$}
                }
                child {node [circle,draw] {} edge from parent[unfixed]
                    child {node [circle,draw,label=right:$x_0^1$] {} edge from parent[unfixed]}
                    edge from parent [fixed] node [right] {$u_2$}
                }
                edge from parent [unfixed]
            };
    \end{tikzpicture}
%%%    \begin{tikzpicture}[overlay]
%%%        \draw [rotate = -20] (-1.9,0.35) ellipse (0.7 and 1.3);
%%%    \end{tikzpicture}
    \caption{A controller-losing game tree}
    \label{fig:interpolatetree}
\end{figure}

Consider the snippet of a game tree in Figure \ref{fig:interpolatetree}. The
tree is losing for the controller, the node labelled $x_1$ is at rank 1, and
$x_0^0$ and $x_0^1$ are at rank 0. Since the tree is controller-losing we know
that at least one run represented by the tree contains the error state.  As
a result, $\textsc{treeFormula}(gt)$ is unsatisfiable. If we take the step
$x_1$ to $x_0^0$ and cut it from the rest of the tree then
$\textsc{treeFormula}(step) \land \textsc{treeFormula}(parent)$ must still be
unsatisfiable.

We can construct an interpolant with $A = \textsc{treeFormula}(parent)$ and $B
= \textsc{treeFormula}(step)$. The only variables shared between $A$ and $B$
are the state variables at $x_1$. We know that $B \land \mathcal{I}$ is
unsatisfiable, therefore all states in $\mathcal{I}$ must lose to the
uncontrollable label $u_1$. We also know that $A \to \mathcal{I}$, thus
$\mathcal{I}$ contains all states reachable by the parent tree (on runs that
avoid the error state.)

\begin{algorithm}
    \caption{Amended tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}
        \Function{treeFormula}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}}$ }
        \Else
        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
        \EndIf
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}
        \Function{$\overline{treeFormula}$}{gt}
        \If{$\Call{rank}{gt} == 0$}
        \State \Return{\Call{E}{$x^{gt}$}}
        \Else
        \State \Return{ $\Call{$L^e[rank(gt)]$}{x^{gt}} \land$ \\
            $(\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n}))$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Now we can consider the step $x_1$ to $x_0^1$, and the parent - now without
either $x_1 \to x_0^0$ or $x_1 \to x_0^1$. The formula
$(\textsc{treeFormula}(parent) \land \mathcal{I}) \land
\textsc{treeFormula}(step)$ must be unsatisfiable. $\mathcal{I}$ contains all
states that lose to $u_1$ so any other state reachable at $x_1$ must lose to
$u_2$. Therefore we can compute another interpolant that contains states that
lose to $u_2$.

This is the foundation for a recursive algorithm that consumes an entire tree
by removing a single step on each iteration. All learned states from which the
controller must lose are recorded in a set $L^c$.  This algorithm can also be
performed on environment-losing trees with the caveat that any state learnt at
a node of rank $n$ is only known to lose at ranks less than or equal to $n$.

\subsection{Proof of Termination}

\subsection{Algorithm}

\section{Evaluation}

\section{Related Work}

\section{Conclusion}
